\section{Introduction}

If artificial intelligence displaces a substantial share of UK employment over the next decade, who pays---the displaced workers, the poorest households, or the Exchequer? The answer depends not only on the size of the shock but also on an assumption that policy analysis often leaves implicit: \emph{who bears it}. The UK's services-intensive economy concentrates employment in the professional, financial, administrative and managerial occupations that score highest on standard measures of AI exposure \citep{felten2021, eloundou2023, pizzinelli2023}, while its safety net---means-tested, tapered, and financed overwhelmingly by taxes on labour income---provides most support through means-tested transfers. A displacement shock aimed at the top of the earnings distribution implicates the welfare state twice over: means-tested support and a labour-income-heavy tax base respond differently depending on which workers lose earnings.

Yet the top-loaded incidence implied by exposure indices is an assumption, not a fact. The empirical literature on AI's labour-market effects has not settled where displacement lands. Exposure-based measures place the risk on high-paid cognitive work \citep{felten2021, eloundou2023}; a competing view holds that AI compresses the returns to accumulated expertise, threatening the wage premia of experienced professionals while opening high-value work to less credentialed workers \citep{autor2024}; and the earliest firm-level evidence finds job losses concentrated among \emph{junior} workers within exposed occupations \citep{kleinteeselink2025, hosseini2026, brynjolfsson2025}. These are not refinements of a single estimate. They are different answers to the incidence question, and they imply different claimants at the benefit office and different holes in the tax base.

This paper treats incidence as an explicit dimension of the scenario design. Rather than committing to one view of who bears the shock, it runs the competing incidence assumptions---not just competing shock sizes---through the same tax and transfer model, translating alternative assumptions about who is displaced into simulated differences in poverty, inequality and Exchequer cost. Using PolicyEngine UK, an open-source microsimulation model running on its processed Family Resources Survey 2024--25 microdata, with AI exposure assigned to Standard Occupational Classification 2020 (SOC2020) occupations \citep{felten2021, tolan2021, eloundou2023}, I simulate a common aggregate shock---seven per cent of employees displaced, an author-applied conversion of quantities discussed by \citet{briggs2023}---under four incidence families: displacement proportional to occupational exposure, displacement concentrated on junior workers, a stylised expertise-compression scenario in which senior premia bear the shock, and a uniform comparator. Around the central calibration I simulate a grid of 66 shock-size combinations crossing displacement rates with wage uplifts, with the capital channel applied throughout; a further robustness family changes the adjustment margin itself, delivering a 7 per cent baseline-wage-bill loss as occupation-level wage cuts with no job loss, graded either by the exposure index or by the potential-supply-shift measure of \citet{hosseinilichtinger2026barriers}. The code and aggregate artefacts are public, but exact regeneration requires licensed FRS microdata; plain survey weights and the absence of the official SPI top-income adjustment also limit representativeness at the top of the distribution. I then examine policy: a fifth, author-constructed top-loaded stress test loosely anchored to UK firm-level evidence, three simulated fiscal responses in the shocked world, a tax-composition accounting of how far corporate profits could offset lost revenue, and a mapping of the shock to the 650 Westminster constituencies.

Two findings organise the paper. First, so long as the shock arrives as job loss, one result is present in every displacement calibration examined: the Gini coefficient of equivalised household disposable income rises in all 66 shock-size cells (by 0.06 to 1.55 percentage points) and in all five incidence families (by 0.93 to 1.13 percentage points at the central shock). Inequality rises regardless of who bears the shock---even though, in the exposure family, the market shock itself is progressive in the sense of falling hardest on the highest deciles. The tax-benefit system claws back part of high earners' losses through forgone tax rather than replacing them through benefits, so a top-loaded market shock still widens the disposable-income distribution.

Second, incidence changes the \emph{form} the simulated cost takes. The same 1.6 million displaced workers cost the Exchequer \pounds 18.2 billion when allocated by exposure (the junior-concentrated mean is similar at \pounds 18.3 billion), \pounds 22.1 billion under expertise compression, and \pounds 14.9 billion under uniform incidence---a spread of \pounds 7 billion a year on an identical aggregate shock, all figures means over 20 displacement draws. Poverty moves less: before-housing-costs poverty rises by 1.75 to 1.93 percentage points across the four stylised families, with no reliable ordering across the 20 assignment draws---while the Klein-anchored stress test exceeds all four on both the fiscal and poverty margins (Section~\ref{sec:results-incidence}). Within these calibrations, top-loaded incidence produces a larger fiscal cost, while flatter and junior-concentrated incidence have smaller mean fiscal effects without correspondingly smaller poverty effects. The adjustment margin matters as much as the allocation: delivering the headline 7 per cent calibration through occupation-level wage cuts rather than job losses---a margin emphasised in equilibrium task models \citep{acemoglu2022tasks}---leaves the fiscal cost comparable (\pounds 22.4 billion) but cuts the poverty rise from +1.87 to +0.06 percentage points and turns the Gini increase into a small decline (Section~\ref{sec:results-wagemargin}). Which world the UK is in cannot yet be settled empirically---so I quantify what is at stake in the disagreement.

The incidence axis matters precisely because exposure indices cannot resolve it. Occupational exposure is age-blind: it cannot distinguish the junior analyst from the senior partner doing related work, so any scenario allocated by exposure alone spreads displacement across the age structure of exposed occupations, which in the UK peaks among prime-age workers. The junior-concentrated pattern in UK firm data \citep{kleinteeselink2025}, US resume histories \citep{hosseini2026} and US payroll records \citep{brynjolfsson2025} must therefore be imposed on the framework rather than derived from it, and the expertise-compression scenario is an author-designed stress test of the mechanism articulated by \citet{autor2024} rather than an estimate.

The remainder of the paper reviews the literature, sets out the scenario and microsimulation methods, reports the distributional and fiscal results, examines three stylised policy responses, and concludes with limitations and implications.
