\section{Related literature}
\label{sec:literature}

My study connects two literatures: competing accounts of where AI-related labour-market adjustment may fall, and microsimulation studies that trace labour-market shocks through tax-benefit systems to household disposable income. I represent the main incidence views as separate scenarios and compare them within the same microsimulation framework.

\subsection{Measuring exposure to artificial intelligence}

The modern exposure literature descends from the occupation-level automation probabilities of \citet{frey2017}, whose estimate that 47 per cent of US employment was at high risk of computerisation set the template of scoring occupations against machine capabilities. Subsequent work replaced whole-occupation probabilities with task- and ability-level measurement: \citet{webb2019} matches patent text to job-task descriptions; \citet{felten2021} link progress in ten AI application areas to the O*NET ability requirements of each occupation, producing the AI Occupational Exposure index I build on; and \citet{tolan2021} connect AI benchmark results to cognitive abilities and tasks. The arrival of large language models prompted a further generation of measures targeted at generative AI specifically: the task-level rubric of \citet{eloundou2023}, the ILO's global assessment of jobs at risk of transformation rather than elimination \citep{ilo2025}, the UK government's occupational assessment \citep{dsit2023}, and the GAISI index of \citet{henseke2025}. What no index can supply is incidence: an index records what AI \emph{could} do to an occupation's task bundle, and translating it into who actually loses work requires assumptions the indices themselves cannot supply. \citet{hosseinilichtinger2026barriers} show that exposure is an incomplete statistic even for the supply side: their measure of the potential supply shift---the expansion of the qualified worker pool as generative AI lowers expertise requirements---varies widely across occupations with identical exposure scores, and they publish it as an alternative public occupation-level measure, which I use as a wage-cut gradient in the wage-margin family of Section~\ref{sec:results-wagemargin}. The complementarity adjustment of \citet{pizzinelli2023}, which I implement in Section~\ref{sec:exposure}, is one such assumption; the incidence families of Section~\ref{sec:incidence} treat the remainder of that translation as an explicit scenario choice.

\subsection{The incidence debate}

Exposure identifies which occupations AI can reach; it does not settle who bears any resulting adjustment. I distinguish three incidence views---exposure-proportional, junior-concentrated and expertise-compression---before reviewing evidence that motivates smaller or differently distributed shocks.

\paragraph{The exposure school.} A common scenario interpretation uses exposure as a proxy for incidence: occupations with higher scores are assigned greater displacement risk and, under that assumption, the indices point towards the top of the labour market. They consistently show exposure rising with education and earnings---professional, managerial and administrative occupations score highest, a pattern that recurs across European labour markets \citep{williamson2025} and, for the UK specifically, in the government's official assessment \citep{dsit2023} and in the GAISI index of \citet{henseke2025}. On this view AI breaks with the routine-biased pattern of earlier waves, in which computerisation hollowed out codifiable mid-skill work \citep{autor2003, goosmanning2007, goos2014, acemoglu2022tasks}, and instead aims at the top of the wage distribution. The IMF strand adds an important qualification: high exposure need not mean displacement, because the same workers may enjoy the largest complementarity gains. \citet{cazzaniga2024} formalise this ambiguity, and \citet{pizzinelli2023} document cross-country variation in the exposure--complementarity mix; \citet{rockall2025} apply the framework to the UK and conclude that AI could \emph{reduce} wage inequality if displacement falls disproportionately on high earners---unless complementarity among those same workers offsets it.

The closest methodological precedent is \citet{doorley2026}. They combine C-AIOE with Irish SILC microdata and the SWITCH tax-benefit model, and their central scenario uses the same 7 per cent employment loss, 2.6 per cent survivor-wage gain and 0.4 percentage-point increase in the return to capital adopted here. Their study provides an Irish exposure-proportional benchmark, alongside uniform-incidence and alternative-index sensitivities. My analysis transfers those allocation mechanics to UK data and extends the design by varying incidence and adjustment margin and by adding UK policy counterfactuals. Differences in country, microdata, occupational resolution and tax-benefit model mean that the numerical results are comparable conditional simulations, not a replication or independent validation. Earlier work by \citet{doorley2023automation} carries automation shocks through tax-benefit models across European countries and finds that progressive taxes and means-tested transfers mute the pass-through of wider wage dispersion to household inequality.

\paragraph{The junior-concentrated evidence.} Several early empirical studies report a junior gradient within exposed occupations. \citet{kleinteeselink2025} link generative-AI exposure to UK firm-level records and find employment reductions of around 4.5 per cent at more exposed firms, concentrated in junior positions; \citet{hosseini2026} document, in US resume and posting data, junior employment falling by around 9 per cent within six quarters of firm AI adoption while senior employment is unaffected; and \citet{brynjolfsson2025} find a 16 per cent relative employment decline among workers aged 22--25 in the most exposed US occupations since late 2022, with older workers largely spared. If incidence is graded by seniority as much as by skill, the distributional consequences differ from a classic skill-biased shock, because junior workers are spread across the household income distribution and interact differently with means-tested benefits.

\paragraph{The expertise-compression school.} A competing view, associated with \citet{autor2024}, holds that AI's distinctive property is not that it targets exposed occupations but that it compresses the returns to accumulated expertise. By supplying decision-relevant knowledge on demand, AI lets less experienced and less credentialed workers perform tasks previously reserved for elite professionals---an opportunity to rebuild middle-class work, but a threat to the wage premia of the experienced workers whose scarcity those premia reflect. Experimental evidence is consistent with the mechanism: generative AI raises output most for the least experienced workers, compressing within-occupation productivity gaps in customer support \citep{brynjolfsson2025qje} and professional writing \citep{noy2023}. \citet{hosseinilichtinger2026barriers} formalise and quantify this mechanism: using LLM ratings of O*NET tasks they construct occupation-level potential supply shifts---the expansion of the qualified worker pool as generative AI lowers expertise requirements---and find in a calibrated general-equilibrium model that entry-barrier erosion compresses between-occupation wage inequality through wage adjustment rather than displacement, with losses concentrated in mid-to-upper-expertise scarcity premia and the extreme top largely insulated. On this view the shock lands on senior premia rather than junior heads---nearly the mirror image of the junior-concentrated evidence above, which is precisely why incidence needs to be varied rather than assumed.

\paragraph{Evidence motivating smaller or differently distributed shocks.} Two further strands inform the scenario range without defining an additional family. The first concerns its \emph{direction}: by making prediction cheap, AI raises the value of the human complements to prediction---judgement, decision-making and responsibility \citep{agrawal2018}---so that workers whose roles centre on decisions rather than predictable task execution may gain, pushing incidence towards prediction-like tasks and away from the senior decision-makers on whom the expertise-compression view lands it. The second concerns its \emph{magnitude}: realised effects may simply be small. Danish population-scale evidence finds minimal effects of AI chatbots on earnings and hours in the first years of diffusion \citep{humlum2025}, US vacancy data show exposed establishments reorganising hiring without detectable aggregate employment effects \citep{acemoglu2022vacancies}, and \citet{acemoglu2025} derives modest macroeconomic gains from task-level foundations. \citet{acemoglu2022tasks} show that even historically transformative automation delivered only a cumulative 3.4 per cent TFP gain over 1980--2016 while reshaping the entire wage structure---large distributional effects do not require large aggregate ones. US evidence on publicly funded job training likewise shows displaced workers from AI-exposed occupations increasingly capturing large retraining returns by sorting towards less-exposed work \citep{hyman2025}. Together these findings motivate varying both incidence and magnitude; my grid therefore extends down to small shocks.

\subsection{From market incomes to disposable incomes}

Whichever school proves right, the welfare consequences run through the tax-benefit system, and market-income incidence is a poor guide to disposable-income incidence. The European evidence on earlier automation makes the point directly: tax-benefit systems absorbed much of the routine-biased shock before it reached household inequality \citep{doorley2023automation}. For AI, the factor-income channel compounds the case for full microsimulation: task-based models imply that automation lowers the labour share and channels income to capital owners \citep{acemoglu2018}, \citet{moll2022} show that automation raises returns to wealth itself, and \citet{korinek2019} develop the theory of worker-replacing progress and the redistribution it calls for---with progressive capital taxation examined as a response by \citet{bastani2024}. Because capital income is far more concentrated than labour income, this channel would tend to raise household inequality, holding the other simulated channels fixed, and only a model that jointly handles earnings, benefits, direct taxes and capital income can net the channels out.

\subsection{Positioning}

Exposure-based tax-benefit simulations exist, while the other incidence views are represented mainly by theory and early empirical estimates. Relative to the exposure-versus-uniform comparison in \citet{doorley2026}, I add junior-concentrated, expertise-compression and Klein-anchored stress-test allocations, hold aggregate displacement fixed across them, and examine adjustment-margin sensitivities. I report how these specified alternatives change simulated poverty, inequality and the public finances.
