\section{Data and methods}
\label{sec:methodology}

The analysis proceeds in three stages. I first attach an occupation-level measure of exposure to artificial intelligence to workers in a nationally grossed UK household survey. I then translate an assumed economy-wide AI shock into person-level changes to employment, wages and capital income, under alternative assumptions about who bears the displacement. Finally, I pass baseline and shocked data through an open-source tax-benefit microsimulation model to recover the distributional and fiscal consequences. This section describes each stage; the incidence scenarios that form the paper's central axis are formalised in Section~\ref{sec:incidence}.

\subsection{Measuring AI exposure}
\label{sec:exposure}

My starting point is the AI Occupational Exposure (AIOE) index of \citet{felten2021}, which links progress in ten AI application areas to the occupational ability requirements catalogued in O*NET, yielding a standardised score for each occupation that captures how much of its ability bundle overlaps with what AI systems can now do. Exposure alone does not distinguish occupations in which AI substitutes for workers from those in which it augments them. I therefore combine the AIOE with the complementarity adjustment of \citet{pizzinelli2023}, who construct an index $\theta_l \in [0,1]$ from O*NET work-context descriptors --- responsibility for others, physical proximity requirements, the consequence of error --- and job-zone preparation requirements. High-$\theta$ occupations are those in which AI is likely to complement human judgement rather than displace it. The complementarity-adjusted exposure measure (C-AIOE) shields high-complementarity occupations from the raw exposure score:
\begin{equation}
\label{eq:caioe}
\text{C-AIOE}_l = \text{AIOE}_l \times \bigl(1 - (\theta_l - \theta_{\min})\bigr),
\end{equation}
where $\theta_{\min}$ is the minimum complementarity across occupational groups, so that the least complementary group retains its full exposure score. Alternative exposure measures include \citet{tolan2021}, the task-level rubric of \citet{eloundou2023}, and the cross-country applications of \citet{cazzaniga2024}; Section~\ref{sec:results} reports the sensitivity of my headline results to substituting two of these for the C-AIOE.

Constructing the index for the UK requires a bespoke crosswalk, whose provenance I state in full because no published UK C-AIOE series exists at the level I need. AIOE scores are taken from the mapping to UK SOC2020 published with the DSIT analysis of AI and UK jobs \citep{dsit2023}, itself derived from the \citet{felten2021} scores via a chained crosswalk (US SOC2018 to US SOC2010, to ISCO-08, to UK SOC2020). The complementarity indices $\theta$ are not published at occupation level, so I reconstruct them from the public O*NET 27.3 release following the recipe documented by \citet{pizzinelli2023}; the reconstruction reproduces the published cross-occupation ordering up to a level offset, which is immaterial because equations~\eqref{eq:caioe} and~\eqref{eq:wage} use $\theta$ only relative to its minimum and its mean. Both series are aggregated to the nine SOC2020 major groups using employment weights from the Annual Survey of Hours and Earnings (ASHE) 2025 Table~14, which covers 340 of the 412 four-digit unit groups; the remainder carry no ASHE employment estimate and are excluded from the weighting. Finally, the C-AIOE scores are shifted so that the least-exposed major group scores exactly zero. Under the allocation rule in Section~\ref{sec:shocks} that group receives no job losses, and the shift prevents negative values of the standardised index from corrupting the allocation weights.

The one-digit level of aggregation is a genuine constraint. The Family Resources Survey (FRS) person records I use carry only the one-digit SOC2020 code, so exposure varies across just nine major groups and all within-group variation is lost. The exposure gradient I impose on the population is correspondingly coarse, and results should be read with that in mind; imputation of finer occupation codes from the Labour Force Survey is a planned refinement.

\subsection{Shock scenarios}
\label{sec:scenarios}

The size of the aggregate shock cannot be estimated from historical data and must be assumed; I treat it as a scenario input and vary it systematically. The central calibration follows the Irish scenario design of \citet{doorley2026}: a displacement rate of 7 per cent of employees and an aggregate wage uplift of 2.6 per cent for those who remain in work. Doorley et al. convert quantities discussed by \citet{briggs2023} into these inputs; they are scenario conventions rather than estimates reported by Briggs and Kodnani or forecasts for the UK. Conceptually, the scenarios calibrate shocks to model primitives---occupation-level automation and productivity, as in the task framework of \citet{acemoglu2022tasks}---whose labour-market consequences the displacement and wage rates then operationalise in reduced form. Alongside the labour-market effects, I follow \citet{doorley2026} in raising the return to capital by 0.4 percentage points, from 1.005 per cent to 1.405 per cent, an increase they motivate using \citet{cazzaniga2024}. \citet{moll2022} supplies only the qualitative mechanism by which automation can raise returns to wealth, not the numerical calibration used here.

Because displacement and wage growth are genuinely uncertain, the central calibration sits inside a grid spanning displacement rates from 0 to 10 per cent and aggregate wage growth from 0 to 5 per cent --- 66 scenario cells in all, with the capital-return shock applied throughout. The 1 per cent ``low'' anchor is a pure sensitivity case: \citet{acemoglu2025} estimates a ten-year GDP effect of around 1 per cent, not an employment displacement rate, and I do not attribute the anchor to him. Beyond the top of the grid, the 13 per cent ``high'' preset stress-tests a displacement rate roughly double the central convention.

\subsection{Implementing the shock at the micro level}
\label{sec:shocks}

\paragraph{Employment shock.} The aggregate number of displaced workers is the scenario displacement rate multiplied by weighted employee employment. Following \citet{doorley2026}, the total is allocated across occupational groups $l$ in proportion to each group's employment multiplied by its exposure. The universe is all employees: workers whose SOC code cannot be matched to an exposure score form a pseudo-group assigned the employment-weighted mean exposure, so no employee is mechanically excluded from risk:
\begin{equation}
\label{eq:jobloss}
\text{JobLoss}_l = \text{TotalJobLoss} \times \frac{\text{EMP}_l \times \text{C-AIOE}_l}{\sum_{l} \text{EMP}_l \times \text{C-AIOE}_l},
\end{equation}
so job losses concentrate in large, highly exposed groups and the zero-scored least-exposed group loses none. Within each group, displaced individuals are selected by systematic sampling on a random permutation of the group's members: the weighted quota from equation~\eqref{eq:jobloss} fixes each record's first-order inclusion probability, which is equal for all records within a group (up to the prescribed probability tilts, such as the youth multiplier, which enter as tilts to these inclusion probabilities), and the systematic pass realises the quota exactly in expectation. A record's inclusion probability therefore does not depend on its grossing weight; the weights enter only through the quota accounting, which is the design intent.

Displacement is modelled as full withdrawal from employee work. A displaced worker's employment income, hours, employee pension contributions and salary-sacrifice amounts are set to zero, any statutory pay is removed, and labour-market status is set to unemployed before the shocked simulation runs. Dual jobholders retain self-employment income. This differs from \citet{doorley2026}, who use a labour-market adjustment add-on to impute unemployment duration and work history for transition cases. The full-year zero-earnings treatment here is an author-chosen UK calibration; the duration sensitivity examines a six-month alternative. Task displacement can instead manifest as wage change and reallocation \citep{acemoglu2022tasks}, and the wage-margin family examines that possibility. These alternatives do not bound all adjustment paths. PolicyEngine UK takes no unemployment-duration input, so displaced workers are assessed as current-period unemployed under standard benefit rules.

\paragraph{Wage shock.} Each worker who survives the employment shock receives a percentage uplift proportional to complementarity, normalised by the employment-weighted mean complementarity over baseline workers:
\begin{equation}
\label{eq:wage}
\text{WageChange}_l = \text{WageGrowth} \times \frac{\theta_l}{\bar{\theta}} \times \text{Wage}_l,
\qquad
\bar{\theta} = \frac{\sum_l \text{EMP}_l \, \theta_l}{\sum_l \text{EMP}_l},
\end{equation}
where $\text{EMP}_l$ is baseline (pre-shock) employment. This is the baseline wage rule. Two properties should be noted. The normalisation makes the \emph{employment-weighted mean} percentage uplift equal the assumed wage growth. The aggregate wage bill grows by exactly that rate only if $\theta$ is uncorrelated with pay. Since high-$\theta$ occupations earn more, the wage-bill-weighted uplift is somewhat larger. A second property: because $\bar{\theta}$ is computed over all baseline workers while the uplift accrues only to survivors---who, being drawn from less-exposed occupations, have above-average $\theta$---displacement raises the realised mean survivor uplift slightly above the headline rate. Exact conservation is verified in seeded tests at zero displacement; both departures are properties of the rule itself rather than implementation error. The economic logic mirrors equation~\eqref{eq:caioe}: AI complements rather than substitutes for high-$\theta$ occupations, so those occupations capture a larger share of the productivity gains --- an assumption the compression scenario of Section~\ref{sec:incidence} deliberately inverts. In the decomposition of \citet{acemoglu2022tasks}, this uplift plays the role of the economy-wide productivity effect, while the displacement effect is carried here entirely by the employment margin; their ripple effects --- displaced workers bidding down the wages of the non-displaced --- are outside the baseline framework, so non-displaced workers can only gain; Appendix~\ref{app:ripple} prices this channel in a reduced-form sensitivity.

\paragraph{Capital shock.} The 0.4 percentage-point increase in the return to capital is implemented by scaling each person's savings-interest and dividend income by the ratio of shocked to baseline return, $1.405/1.005 \approx 1.398$. The baseline return of 1.005 per cent is taken from the baseline calibration; because the scaling factor is the ratio of two return levels, the size of the capital channel is highly sensitive to this assumption. The same $+0.4$ percentage-point shock applied to, say, a 4 per cent baseline return would scale capital incomes by only 10 per cent rather than 40 per cent. The capital-channel results should therefore be read as conditional on this calibration; equivalently, the shock can be read directly as a 39.8 per cent proportional increase in interest and dividend income. In the framework of \citet{moll2022} the relevant return is that on investors' wealth rather than the safe rate, so a refinement would load the shock onto dividend income alone; loading the shock on dividends alone could produce a more top-concentrated incidence, but I do not quantify that alternative. Rental income is excluded from the scaling, on the modelling assumption that the AI productivity channel operates through financial rather than housing capital. The self-employed are outside both labour-market shocks; their interest and dividend income is scaled like everyone else's.

\subsection{Incidence scenarios}
\label{sec:incidence}

The rules above fix how large the shock is; they do not settle who bears it, and the empirical literature has not settled that question either. I therefore treat incidence as a explicit scenario dimension. Holding the central aggregate shock fixed--- 7 per cent displacement, a 2.6 per cent wage uplift, the capital shock--- I re-allocate the same weighted quota of displacement under four incidence families:

\paragraph{Exposure-proportional.} The default allocation of equations~\eqref{eq:jobloss} and~\eqref{eq:wage}: displacement follows the C-AIOE gradient and wage gains follow complementarity. This family is the central calibration used throughout the shock-size grid.

\paragraph{Junior-concentrated.} Displacement draws are tilted towards workers aged 16--24 by a multiplier equal to the ratio of junior to overall displacement effects reported by \citet{kleinteeselink2025} for the UK ($5.8/4.5$), consistent with the seniority-biased evidence of \citet{hosseini2026} and \citet{brynjolfsson2025}. The group quotas and wage rule are otherwise unchanged.

\paragraph{Expertise-compression.} A stylised, author-designed stress test of the view that AI compresses expert rents rather than displacing routine work; as implemented it is a \emph{high-earner displacement tilt}, and the label ``expertise-compression'' names the motivating hypothesis rather than a modelled compression mechanism. Displacement probabilities are tilted (multiplier 2, an author choice) towards the top weighted-earnings tertile of workers in above-median-exposure occupations. The survivor-wage and capital channels remain exactly those of the exposure-proportional family, so the headline factorial comparison changes only who is displaced. An older compound variant that also reflects $\theta$ about its range is retained in the committed outputs under the explicit \texttt{compression\_compound} label but is not part of the headline comparison. The formal treatment in \citet{hosseinilichtinger2026barriers} operates through wage compression for continuing workers rather than job destruction; this displacement-based implementation is therefore a fiscal stress test, not an implementation of their model.

\paragraph{Uniform.} Every employee has the same first-order displacement probability, while the central complementarity-graded survivor-wage channel is held fixed. Relative to exposure-proportional incidence, it therefore isolates a flat displacement allocation. The earlier bundle that flattened both displacement and wages remains available only as \texttt{uniform\_compound}.

Each displacement family has the same expected weighted count of displaced workers. Cross-family differences in the headline factorial comparison reflect displacement incidence alone; survivor wages and capital income are held fixed at the person level. These rules are scenario assumptions, not estimates of realised AI displacement.

\paragraph{Klein-anchored stress test.} Section~\ref{sec:results-incidence} adds a fifth, explicitly author-constructed family loosely anchored to the UK firm-level evidence of \citet{kleinteeselink2025} (named \texttt{klein\_top\_loaded} in the committed artefacts). Displacement probability is proportional to shifted C-AIOE multiplied by a junior factor of 1.29 and wage-tier factors of 9.6 above versus 0.6 below weighted-median person earnings; the aggregate shock stays at the central preset. These values are not a transported estimate of household-level displacement. The revised paper reports per-standard-deviation semi-elasticities of about $-0.3$ per cent overall, $-0.4$ per cent for junior employment, and $-0.7$ per cent in high-compensation firms against a statistical zero in low-compensation firms. The much larger wage-tier factors used here descend from press-release scalings to maximally exposed firms, and person earnings only proxy the paper's firm-average compensation split. I remove the previously imposed London multiplier because the December 2025 revision reports minimal, statistically insignificant London effects. The remaining junior proxy (age under 25), wage-tier mapping and multiplicative combination are author choices, so this family is a top-loaded stress test, not a measured wage or displacement schedule.

\paragraph{Wage margin.} The families above vary who is displaced; a final family varies the \emph{margin} on which the shock arrives, and I treat it as a separate adjustment-margin robustness family rather than a sixth incidence family, since it changes the mechanism rather than the allocation. In the task framework of \citet{acemoglu2022tasks}, displacement ``does not need to be associated with job loss'' and manifests chiefly as relative wage declines for workers who stay employed; the entry-barrier mechanism of \citet{hosseinilichtinger2026barriers} likewise operates through wage adjustment rather than job destruction. The wage-margin family moves the same aggregate earnings loss to that channel: no worker loses a job, and instead every employee's earnings are cut by an occupation-level percentage proportional to $\max(0, \text{exposure} - \text{exposure}_{\min})$ --- the same min-shift normalisation as the quota weights of equation~\eqref{eq:jobloss}, so the least-exposed group takes no cut---with a single calibration constant chosen so that the weighted aggregate employee-earnings loss equals the \emph{realised} gross earnings loss of the paired central displacement draw for the same seed (a realised paired share of 7.77 per cent of the baseline employee wage bill at seed 0---0.07774, recorded with the output---larger than the 7 per cent head-count rate because group job-loss rates correlate with group pay; the equality is asserted in code). The family therefore shares its calibration draw-by-draw with the central scenario rather than with a fixed 7 per cent wage-bill share, and its only seed variation enters through the paired draw's realised loss. % results/wage_margin_central.json gross_cut_share_target
I run two gradients: the C-AIOE itself, and the potential-supply-shift (PSS) measure of \citet{hosseinilichtinger2026barriers}, taken from their public data release and crosswalked to SOC2020 major groups by the same chained SOC2018--SOC2010--ISCO-08--SOC2020 route with ASHE 2025 employment weights as the exposure scores (Spearman rank correlation with the C-AIOE across the nine major groups: 0.78). The PSS rows should be read as a \emph{PSS-ranked wage-cut stress test}: PSS supplies only the cross-occupation ranking of author-imposed cuts, and the exercise is not an implementation of the \citet{hosseinilichtinger2026barriers} labour-supply mechanism, which their citation here motivates rather than licenses. The 2.6 per cent complementarity-graded wage uplift of equation~\eqref{eq:wage} and the capital shock compose on top, so a worker's net percentage change is uplift minus cut; employment status, hours and pension contributions are unchanged. Section~\ref{sec:results-wagemargin} reports the results.

\paragraph{Mixed adjustment and adverse wage pass-through.} Two further sensitivities relax the assumption that adjustment occurs entirely on one margin and that workers who retain their jobs necessarily gain. First, for each seed I calculate the gross pre-uplift earnings removed by the central 7 per cent employee-displacement draw and hold that amount fixed. Let $\lambda\in\{0,.25,.5,.75,1\}$ scale the employee-displacement rate, $0.07\lambda$. C-AIOE-graded cuts among survivors then fill the exact gap between earnings removed through partial displacement and the seed-specific central gross-loss target. The usual 2.6 per cent complementarity-graded uplift and capital shock compose only after this identity is imposed. Thus $\lambda=1$ exactly reproduces the central displacement scenario for the same seed; $\lambda=0$ is a pure wage-cut counterfactual matched to the central draw's realised loss, not the separate family that removes exactly 7 per cent of the baseline wage bill. The intermediate cells are author-designed reduced-form comparative statics rather than equilibrium transition paths. Second, holding the central 7 per cent displacement draw fixed, I replace the assumed average survivor-wage effect with $\{-5,-2.6,0,+2.6,+5\}$ per cent, retaining the complementarity gradient. This grid includes wage suppression, no pass-through, the central calibration and a larger productivity pass-through without requiring a model of firm wage-setting or bargaining.

\subsection{Microsimulation with PolicyEngine UK}
\label{sec:microsim}

I use PolicyEngine UK (version 2.89.2), an open-source tax-benefit microsimulation model, running on PolicyEngine UK's processed (base, unenhanced) Family Resources Survey 2024--25 microdataset. This reshapes the UK Data Service FRS 2024--25 (study SN~9563) into PolicyEngine's tax-benefit schema, applies source-level imputations (council tax, benefit take-up, salary sacrifice) and retains the survey's own grossing weights; I obtain it from PolicyEngine's \texttt{policyengine-uk-data} repository and load it through \texttt{UKSingleYearDataset} and \texttt{Microsimulation}. Occupation codes are not carried in the processed dataset, so I join the SOC2020 major group from the raw FRS \texttt{adult.tab} file (UK Data Service SN~9563) onto the simulation's person table using the identifier convention $\text{person\_id} = \text{SERNUM} \times 1000 + \text{PERSON}$, which matches every FRS adult. The simulation year is 2026.

The counterfactual design is a paired comparison on identical data: a baseline simulation and a shocked simulation are run on the same records, with the shocked run receiving the modified employment income, hours, pension contributions, savings-interest income, dividend income and employment status of Section~\ref{sec:shocks} as direct inputs. All reported effects are differences between the two runs. This design removes between-run noise --- the survey sample and grossing weights are identical on both sides, so no part of a reported effect reflects resampling --- but it does not quantify population sampling uncertainty: both runs inherit whatever sampling error the FRS itself carries, and my Monte Carlo intervals speak only to the stochastic identity of the displaced, not to survey sampling.

All income results use the HBAI cash disposable income concept. Absolute BHC and AHC poverty compare equivalised household income with fixed real thresholds derived from the baseline; those thresholds are frozen in baseline--shock comparisons. Relative BHC and AHC poverty instead use 60 per cent of the contemporaneous median equivalised income, so their thresholds move with the simulated distribution. I report the government-balance change; person-weighted BHC and AHC poverty; the Gini of equivalised household disposable income, with negative incomes bottom-coded at zero; and mean household-income changes by baseline-fixed decile.

Because the identity of the displaced is stochastic, I adopt explicit seed conventions and label every result accordingly. The 66-cell grid, the three presets and the decomposition figure are single draws at seed~0, so that cells differ only in their scenario inputs, never in their random draw. The five incidence families report means over 20 draws (seeds 0--19). The decile transition and wage profiles (Figures~\ref{fig:transition} and~\ref{fig:wages}) average 50 seeded draws. The robustness summary and the decomposition assignment-draw band use a 20-draw Monte Carlo at seeds 0--19; wherever a $\pm$ value appears it is the standard deviation across draws, not a survey-based standard error.

\paragraph{Policy reforms in the shocked world.}The reforms of Section~\ref{sec:policy} are evaluated as $M(\text{shock}+\text{reform}) - M(\text{shock})$ on the fixed seed-0 shock draw, applying for 2026 only, with poverty lines held at their shocked-world values. Benefit-side reforms are implemented as PolicyEngine UK parameter reforms --- scaling the Universal Credit (UC) standard allowance, raising the benefit-cap thresholds and cutting the taper rate --- so their costs are net of the means-testing clawback the model computes. Wage insurance is instead a direct post-simulation transfer to displaced workers (half of lost earnings, capped), treated as non-taxable and disregarded by means tests, so its gross cost equals its net cost and brackets the generous end of possible implementations.

\paragraph{Constituency geography.}The constituency results of Section~\ref{sec:results-geo} rest on PolicyEngine's $650 \times 53{,}508$ weight matrix, which assigns each household in the \emph{enhanced} FRS 2023--24 dataset a calibrated weight in every Westminster parliamentary constituency. The SERNUM-based occupation join of Section~\ref{sec:microsim} is invalid on that dataset, so SOC major group is imputed from the plain-FRS 2024--25 weighted distribution within age band $\times$ gender $\times$ region $\times$ earnings-decile cells (98.9 per cent weighted coverage from the full four-way cell, seed 0). % results/geo/imputation_notes.json
Constituency variation therefore reflects demographic and earnings composition plus a single displacement draw, not observed occupation mix, and the geographic results --- computed on a different dataset and period --- are not directly comparable to the poverty numbers elsewhere in the paper.
