\documentclass[11pt]{article}
\usepackage[margin=1.1in]{geometry}
\usepackage{booktabs}
\usepackage{amsmath}
\usepackage[hidelinks]{hyperref}
\usepackage{natbib}
\bibliographystyle{apalike}

\title{Artificial Intelligence and Income Inequality in the United Kingdom:\\
\large A replication of Doorley et al.\ (2026) with PolicyEngine UK\thanks{Draft v0.1. Code and aggregate results: \url{https://github.com/PolicyEngine/uk-ai-study}. FRS microdata via the UK Data Service (EUL); not redistributed.}}
\author{Vahid Ahmadi}
\date{July 2026 --- preliminary draft}

\begin{document}
\maketitle

\begin{abstract}
\noindent We replicate for the United Kingdom the scenario framework of
Doorley, O'Connor, O'Shea and Tuda (2026, ESRI/Department of Finance Report
No.~16), which traces generative AI's labour-market effects through a
tax-benefit microsimulation. Occupation-level AI exposure
(complementarity-adjusted AIOE) is attached to Family Resources Survey
adults; employment, wage and capital shocks follow the report's equations
3.4 and 3.5; and fiscal and distributional outcomes are computed with
PolicyEngine UK in place of SWITCH. In the central scenario (7\%
displacement, +2.6\% wages, +0.4pp return to capital), the annual Exchequer
cost is \pounds 41bn, the before-housing-costs poverty rate rises by
3.3 percentage points, and the Gini of equivalised household net income
rises by 0.019. We extend the framework along the age dimension: tilting
within-occupation displacement toward workers aged 16--24, per recent
seniority-biased evidence, raises the youth share of the displaced by
roughly two-fifths while leaving aggregate effects nearly unchanged.
\end{abstract}

\section{Introduction}

Doorley et al.\ (2026) ask how generative AI would affect income inequality
in Ireland if it displaces some workers, raises the wages of those who
remain, and raises returns to capital. Their answer runs occupation-level
AI-exposure scores through stylised shock scenarios and the SWITCH
tax-benefit model. We port the full pipeline to the UK: Family Resources
Survey (FRS) microdata replaces SILC, SOC2020 occupations replace ISCO, and
PolicyEngine UK --- an open-source tax-benefit microsimulation model ---
replaces SWITCH. We add one substantive extension: an age dimension,
motivated by evidence that generative AI's employment effects concentrate
among junior workers \citep{kleinteeselink2025, hosseini2026,
brynjolfsson2025}.

\section{Data and method}

\subsection{Exposure}

Each FRS 2024--25 adult reports a SOC2020 major group (\texttt{adult.tab},
UK Data Service). We attach the complementarity-adjusted AI Occupational
Exposure score (C-AIOE) of \citet{pizzinelli2023} --- the \citet{felten2021}
AIOE scaled by an O*NET-derived complementarity index $\theta$ --- using a
UK SOC2020 crosswalk built from open-licensed sources (Felten et al.'s
published scores, a reconstruction of $\theta$ from public O*NET 27.3
following the IMF recipe, and ASHE 2025 employment weights for the
major-group aggregation). Version 0.1 operates at the 1-digit major-group
level; within-group exposure variation is deferred to a planned
quantile-forest imputation from 4-digit LFS occupations.

\subsection{Shocks}

Following the report's equation~3.4, aggregate displacement equals the
scenario rate times weighted employment; it is allocated across major groups
in proportion to employment times mean C-AIOE and realised by weighted
random draws within groups (probability proportional to individual exposure).
Displaced workers receive zero employment income and unemployed status.
Following equation~3.5, a wage pool equal to the scenario uplift times the
surviving wage bill is distributed across survivors in proportion to
$\theta$ times earnings. Interest and dividend income are scaled by the
ratio of shocked to baseline capital returns ($1.405\%/1.005\% \approx
1.398$). Scenario anchors: central 7\% displacement with +2.6\% wages
\citep{briggs2023}; low 1\% employment-only \citep{acemoglu2025}; high 13\%
\citep{brynjolfsson2025}. The youth-tilted variant multiplies the draw
probability of workers aged 16--24 by $5.8/4.5 \approx 1.29$, the ratio of
junior to total employment effects in \citet{kleinteeselink2025}, holding
group totals fixed.

\subsection{Microsimulation}

Baseline and shocked PolicyEngine UK simulations (v2.89.2, simulation year
2026) run on the same FRS dataset; the shocked run receives modified
employment, interest and dividend income and employment status. We report
the change in government balance (Exchequer cost), person-weighted poverty
rates before and after housing costs, and the Gini of equivalised household
net income, with breakdowns by baseline income decile and age band.

\section{Results}

\begin{table}[h]
\centering
\caption{Scenario results, FRS 2024--25, simulation year 2026}
\begin{tabular}{lrrrr}
\toprule
 & Displacement & Exchequer cost & $\Delta$ Poverty (BHC) & $\Delta$ Gini \\
\midrule
Low & 1\% & \pounds 6.5bn & +0.32pp & +0.003 \\
Central & 7\% & \pounds 41.1bn & +3.25pp & +0.019 \\
High & 13\% & \pounds 93.9bn & +7.02pp & +0.034 \\
Central, youth-tilted & 7\% & \pounds 39.4bn & +3.42pp & +0.018 \\
\bottomrule
\end{tabular}
\end{table}

Fiscal costs scale roughly linearly at \pounds 6--7bn per percentage point
of displacement, on top of an always-on capital-shock component. The
youth-tilted central scenario raises the 16--24 share of the displaced from
3.2\% to 4.5\% --- a two-fifths increase --- while aggregate fiscal and
inequality effects move little, because young workers' earnings and benefit
entitlements are smaller than average. Poverty, however, rises slightly
\emph{more} under the youth tilt (+3.42pp vs +3.25pp).

\section{Limitations}

Exposure is 1-digit only; weights are the uncalibrated FRS build; displaced
workers are current-period unemployed rather than the report's long-term
unemployed contract (benefit-exhaustion nuances are not expressible as
PolicyEngine inputs); the self-employed sit outside all shocks, as in the
original; and results reflect a single seeded draw per scenario. Each is
addressed in the repository roadmap.

\begin{thebibliography}{9}
\bibitem[Acemoglu(2025)]{acemoglu2025} Acemoglu, D. (2025). The simple macroeconomics of AI. \emph{Economic Policy}, 40(121).
\bibitem[Briggs \& Kodnani(2023)]{briggs2023} Briggs, J. \& Kodnani, D. (2023). The potentially large effects of artificial intelligence on economic growth. Goldman Sachs.
\bibitem[Brynjolfsson et al.(2025)]{brynjolfsson2025} Brynjolfsson, E., Chandar, B. \& Chen, R. (2025). Canaries in the coal mine? Six facts about the recent employment effects of artificial intelligence.
\bibitem[Doorley et al.(2026)]{doorley2026} Doorley, K., O'Connor, S., O'Shea, R. \& Tuda, D. (2026). \emph{Artificial intelligence and income inequality in Ireland}. ESRI/Department of Finance Report No.~16.
\bibitem[Felten et al.(2021)]{felten2021} Felten, E., Raj, M. \& Seamans, R. (2021). Occupational, industry, and geographic exposure to artificial intelligence. \emph{Strategic Management Journal}, 42(12).
\bibitem[Hosseini \& Lichtinger(2026)]{hosseini2026} Hosseini, S.M. \& Lichtinger, G. (2026). Generative AI as seniority-biased technological change. SSRN 5425555.
\bibitem[Klein Teeselink(2025)]{kleinteeselink2025} Klein Teeselink, B. (2025). Generative AI and labor market outcomes: Evidence from the United Kingdom. SSRN 5516798.
\bibitem[Pizzinelli et al.(2023)]{pizzinelli2023} Pizzinelli, C., Panton, A., Tavares, M., Cazzaniga, M. \& Li, L. (2023). Labor market exposure to AI: Cross-country differences and distributional implications. IMF WP/23/216.
\end{thebibliography}

\end{document}
